Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex
Updated

\section{Experiments}\label{sec:exp}
In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer. This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}. Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$. In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates. Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration. To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}). Each experiment is run 10 times with a random policy initialization, different each time, shared by all the algorithms under comparison. The length of the experiment, \ie the total number of trajectories, is fixed for each task. Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process. We provide average performance with 90\% bootstrap confidence intervals.  
Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based. More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.

@@

Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task. Despite using more trajectories for each parameter, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.

@@

Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning~\citep[\eg][]{duan2016benchmarking}. This corresponds to the area below the curves in the figures. Note also that the convergence result Theorem \ref{theo:convergence} is precisely about average return. We also report normalized standard deviation and 90\% bootstrap confidence intervals.
In all the tasks, our algorithm has higher average return and lower normalized standard deviation. In the case of Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.

\paragraph{Preliminary results on actor-critic.}As mentioned, another variance-reduction technique in policy gradient is to employ baselines, or \textit{critics}. This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense. In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without baseline. In order to move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario. To do so, we use the more challenging mujoco Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer. Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms. Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.

>>>>

\section{Experiments}\label{sec:exp}
In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer.\todopir{ref, even Benchmarking Deep Reinforcement Learning for Continuous Control is fine}
We consider G(PO)MDP since has a smaller variance than REINFORCE.
% This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. 
For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}. Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$. In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates. Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration. To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
Each experiment is run $10$ times with a random policy initialization and seed but this initialization is shared among the algorithms under comparison.
The length of the experiment, \ie the total number of trajectories, is fixed for each task. Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process. We provide average performance with 90\% bootstrap confidence intervals.  
Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{rllab can be found at \url{https://github.com/rll/rllab}, while the link to our implementation is not reported for blindness but it is already publically available.}
More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.

@@

Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task. Despite using more trajectories for each parameter,\todopirout{I do not understand this sentence} our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.

@@

Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}. This corresponds to the area below the curves in the figures. Note that the convergence result Theorem \ref{theo:convergence} is precisely about average return. We also report normalized standard deviation and 90\% bootstrap confidence intervals.
In all the tasks, our algorithm has higher average return and lower normalized standard deviation.
In Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.

\textbf{Preliminary results on actor-critic.} %As mentioned, 
Another variance-reduction technique in policy gradient is to employ baselines, or \textit{critics}. This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense. In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without baseline. In order to move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario. To do so, we use the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer. Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms. Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
